---
title: "Inhibitory cells post N49"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

take all inh cells used in perprint, and add nova49 results.


```{r}
library(Seurat)
library(scrattch.hicat)
library(ggplot2)
library(dplyr)
library(patchwork)
```

load all neurons

```{r,fig.width=15,fig.height=5}
setwd('~/postdoc2/DCN_sequencing/human/plateseq/novaseq46/')
load('neurons.RData')
p1<-DimPlot(neurons,label=T)
p2<-FeaturePlot(neurons,c("SLC17A6","GAD1"))
(p1 | p2)
```



annotate by FACS round:
```{r}
plates=sapply(strsplit(rownames(neurons@meta.data),"_"),function(x) x[1])
R01=plates %in% c("LLL29","LLL30")
R02=plates %in% c("LLL31","LLL32","LLL33")

R1=plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL51")
R2=plates %in% c("LLL50")
R3=plates %in% c("LLL51","LLL61","LLL53","LLL59","LLL60","LLL63")
R4=plates %in% c("LLL54","LLL55","LLL56")
R5=plates %in% c("LLL57","LLL58")
R6=plates %in% c("LLL79","LLL83")
R7=plates %in% c("LLL80","LLL81","LLL82","LLL84")

FACS=as.factor(as.numeric(R01)+2*as.numeric(R02)+3*as.numeric(R1)+4*as.numeric(R2)+5*as.numeric(R3)+6*as.numeric(R4)+7*as.numeric(R5)+8*as.numeric(R6)+9*as.numeric(R7))
neurons$FACS<-FACS
```



```{r}
inh1<-subset(neurons,idents=c(1,2,4,6,8,10,9))
inh1<-subset(inh1,subset= SLC17A6<4)
load('~/postdoc2/DCN_sequencing/human/plateseq/novaseq46/contaminatingGolgicells.RData')
inh1<-subset(inh1,cells=golgi, invert=T)

```


```{r,fig.width=15,fig.height=5}
setwd('~/postdoc2/DCN_sequencing/human/plateseq/novaseq49/')
load('neurons.RData')
inh2<-subset(neurons,idents=c(0,8,7,3))
```

annotate by FACS round:
```{r}
plates=sapply(strsplit(rownames(inh2@meta.data),"_"),function(x) x[1])
R1=plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL51")
R2=plates %in% c("LLL50")
R3=plates %in% c("LLL51","LLL61","LLL53","LLL59","LLL60","LLL63")
R4=plates %in% c("LLL54","LLL55","LLL56")
R5=plates %in% c("LLL57","LLL58")
R6=plates %in% c()
R7=plates %in% c()
R8=plates %in% c("LLL90","LLL91","LLL92","LLL93","LLL94")
R9=plates %in% c("LLL96","LLL97","LLL98","LLL99","LLL100","LLL101","LLL102")
FACS=as.factor(as.numeric(R1)+2*as.numeric(R2)+3*as.numeric(R3)+4*as.numeric(R4)+5*as.numeric(R5)+6*as.numeric(R6)+7*as.numeric(R7)+8*as.numeric(R8)+9*as.numeric(R9))
inh2$FACS<-FACS
```






```{r}
inh<-merge(inh1,inh2)
```

drop MT- genes
```{r}
mt<-grep("^MT-",rownames(inh),invert = T)
inh<-inh[mt,]
```

```{r}
inh <- FindVariableFeatures(object = inh,selection.method = "vst",nfeatures = 2000,verbose=FALSE)
inh <- ScaleData(object = inh,vars.to.regress = c('nFeature_RNA','donor'))
#inh <- RunPCA(object = inh,npcs = 30,verbose = FALSE)


source('~/postdoc2/DCN_sequencing/scRNAseq/plateseq/analysis_afterNova15/crossDCNalignment/pc_modification_functions_S3_forRNA.R')

inh<-RunTruncatedPCA(inh,n.genes.pc = 60)
ElbowPlot(object = inh)
```
```{r, fig.height=6, fig.width=10, warning=FALSE}
#inh <- JackStraw(object = inh, dims=20)
#inh <- ScoreJackStraw(inh,dims=1:20)
#JackStrawPlot(object = inh,dims=1:20)
```

```{r}
usefuldims=1:10
dims.remove=c()
usefuldims=usefuldims[!usefuldims %in% dims.remove]
```

```{r, fig.height=5, fig.width=7}
inh<- FindNeighbors(inh,dims=usefuldims)
inh <- FindClusters(object = inh, resolution = 2)
```



```{r, fig.height=10, fig.width=15, message=FALSE, warnings=FALSE}
inh <- RunTSNE(object = inh, dims = usefuldims, perplinhity=30, dim.embed = 2)


p1<-DimPlot(object = inh, reduction = 'tsne', label=TRUE,pt.size = 1.5)
p2<-DimPlot(inh,group.by = 'donor',pt.size = 1.5)
p3<-DimPlot(inh,group.by = 'DCN',pt.size = 1.5)
p4<-DimPlot(inh,group.by = 'orig.ident',pt.size = 1.5)
(p1 | p2)/(p3 | p4)
```

merge clusters using allen package.
```{r}
library(scrattch.hicat)
rd.dat <- t(GetAssayData(object = inh, slot = "scale.data"))

merge.param <- de_param(padj.th     = 0.05, 
                     lfc.th      = 1, 
                     low.th      = 1, 
                     q1.th       = 0.4, 
                     q.diff.th   = 0.6, 
                     de.score.th = 40)

merge.result <- merge_cl(as.matrix(GetAssayData(object = inh, slot = "data")), 
                         cl = inh$RNA_snn_res.2, 
                         rd.dat = rd.dat,
                         de.param = merge.param)


if (is.null(merge.result))
  {inh$merged.res.2<-'inh_1'
} else {
inh$merged.res.2<-as.factor(paste('inh_',merge.result$cl,sep=''))
}
Idents(inh)<-'merged.res.2'
```

```{r,fig.width=15,fig.height=5}
p1<-DimPlot(inh,label=T,group.by = 'merged.res.2',pt.size = 1.5)
p2<-DimPlot(inh,label=T,group.by = 'RNA_snn_res.2',pt.size=1.5)
(p1 | p2)
```



```{r}
VlnPlot(inh,c("nFeature_RNA","nCount_RNA"),group.by = 'merged.res.2')
```


```{r}
FeaturePlot(inh,c("SNAP25","GAD1","SLC17A6","SLC6A5"))
```

Some donor effects.

Let's try to tackle donor effects by integrating inh data by donor for all donors

```{r}
options(future.globals.maxSize= 891289600)

Idents(inh)<-'donor'
inh_3donors<-subset(inh,idents=c('B0','B1','B3'))
CN.list <- SplitObject(inh_3donors, split.by = "donor")
for (i in 1:length(CN.list)) {
    CN.list[[i]] <- NormalizeData(CN.list[[i]], verbose = FALSE,scale.factor = 1e6)
    CN.list[[i]] <- FindVariableFeatures(CN.list[[i]], selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
}
CN.anchors <- FindIntegrationAnchors(object.list = CN.list, dims = 1:30,k.filter = 100)
CN.integrated <- IntegrateData(anchorset = CN.anchors, dims = 1:30,features.to.integrate = rownames(inh_3donors))

```
```{r}
DefaultAssay(CN.integrated) <- "integrated"
CN.integrated <- ScaleData(CN.integrated, verbose = FALSE,vars.to.regress = c("FACS",'nCount_RNA'))
CN.integrated <- RunPCA(CN.integrated, npcs = 30, verbose = FALSE)
ElbowPlot(CN.integrated)
```



```{r}
usefuldims=1:5
dims.remove=c()
usefuldims=usefuldims[!usefuldims %in% dims.remove]
```

```{r, fig.height=5, fig.width=7}
CN.integrated<- FindNeighbors(CN.integrated,dims=usefuldims)
CN.integrated <- FindClusters(object = CN.integrated, resolution = 2)
```
```{r, fig.height=10, fig.width=15, message=FALSE, warnings=FALSE}
CN.integrated <- RunTSNE(object = CN.integrated, dims = usefuldims, perplinhity=20, dim.embed = 2)


p1<-DimPlot(object = CN.integrated, reduction = 'tsne', label=TRUE,pt.size = 1.5,group.by = 'integrated_snn_res.1')
p2<-DimPlot(CN.integrated,group.by = 'donor',pt.size = 1.5)
p3<-DimPlot(CN.integrated,group.by = 'DCN',pt.size = 1.5)
p4<-DimPlot(CN.integrated,group.by = 'orig.ident',pt.size = 1.5)
(p1 | p2)/(p3 | p4)
```

merge clusters using allen package.
```{r}
library(scrattch.hicat)
rd.dat <- t(GetAssayData(object = CN.integrated, slot = "scale.data"))

merge.param <- de_param(padj.th     = 0.05, 
                     lfc.th      = 1, 
                     low.th      = 1, 
                     q1.th       = 0.4, 
                     q.diff.th   = 0.6, 
                     de.score.th = 40)

merge.result <- merge_cl(as.matrix(GetAssayData(object = CN.integrated, slot = "data",assay = 'integrated')), 
                         cl = CN.integrated$integrated_snn_res.2, 
                         rd.dat = rd.dat,
                         de.param = merge.param)


if (is.null(merge.result))
  {CN.integrated$merged.res.2<-'inh_1'
} else {
CN.integrated$merged.res.2<-as.factor(paste('inh_',merge.result$cl,sep=''))
}
Idents(CN.integrated)<-'merged.res.2'
```

```{r,fig.width=15,fig.height=5}
p1<-DimPlot(CN.integrated,label=T,group.by = 'merged.res.2',pt.size = 1.5)
p2<-DimPlot(CN.integrated,label=T,group.by = 'DCN',pt.size=1.5)
(p1 | p2)
```
```{r}
VlnPlot(CN.integrated,c("SLC6A5","GAD1","SLC17A6","SNAP25"),assay = 'RNA')
```

```{r,fig.width=15,fig.height=20}
FeaturePlot(CN.integrated,paste0("PC_",1:20),cols = c("navy", "olivedrab1"),pt.size = 1)
```
```{r}
FeaturePlot(CN.integrated,"rna_SLC6A5",pt.size=1.5)
```


```{r}
#save(CN.integrated,file = "~/postdoc2/DCN_sequencing/human/plateseq/Novaseq49/inh_3donors_integrated.RData")
```


i1.2 maps to CL11 in the mouse dataset, and intermingles with excitatory cells. It's also very rare <20 cells per donor, 32 total out of 2413. so lets eliminated it.

```{r}
Idents(CN.integrated)<-'merged.res.2'
CN.integrated<-RenameIdents(CN.integrated,'inh_3'='i1.1','inh_18'='i1.2','inh_1'='i2.1','inh_16'='i2.2','inh_0'='i2.3','inh_7'='i3')

CN.integrated$merged.res.2.renamed<-Idents(CN.integrated)
CN.integrated<-subset(CN.integrated,idents='i1.2',invert=T)
save(CN.integrated,file = "~/postdoc2/DCN_sequencing/human/plateseq/Novaseq49/inh_clean_3donors_integrated.RData")
DimPlot(CN.integrated)
```




