---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 
```{r}
library(Seurat)
library(dplyr)
library(cowplot)
library(ggplot2)
library(patchwork)
library(future)
plan("multiprocess",workers=4)
```







load Novaseq49 data
```{r}

setwd('~/postdoc2/DCN_sequencing/human/plateseq/novaseq49/')

raw_counts90e<-read.table(file="genesXcells_exons_LLL90.tab",sep="\t",row.names=1,header=TRUE)
raw_counts91e<-read.table(file="genesXcells_exons_LLL91.tab",sep="\t",row.names=1,header=TRUE)
raw_counts92e<-read.table(file="genesXcells_exons_LLL92.tab",sep="\t",row.names=1,header=TRUE)
raw_counts93e<-read.table(file="genesXcells_exons_LLL93.tab",sep="\t",row.names=1,header=TRUE)
raw_counts94e<-read.table(file="genesXcells_exons_LLL94.tab",sep="\t",row.names=1,header=TRUE)
raw_counts96e<-read.table(file="genesXcells_exons_LLL96.tab",sep="\t",row.names=1,header=TRUE)
raw_counts97e<-read.table(file="genesXcells_exons_LLL97.tab",sep="\t",row.names=1,header=TRUE)
raw_counts98e<-read.table(file="genesXcells_exons_LLL98.tab",sep="\t",row.names=1,header=TRUE)
raw_counts99e<-read.table(file="genesXcells_exons_LLL99.tab",sep="\t",row.names=1,header=TRUE)
raw_counts100e<-read.table(file="genesXcells_exons_LLL100.tab",sep="\t",row.names=1,header=TRUE)
raw_counts101e<-read.table(file="genesXcells_exons_LLL101.tab",sep="\t",row.names=1,header=TRUE)
raw_counts102e<-read.table(file="genesXcells_exons_LLL102.tab",sep="\t",row.names=1,header=TRUE)

raw_counts90i<-read.table(file="genesXcells_introns_LLL90.tab",sep="\t",row.names=1,header=TRUE)
raw_counts91i<-read.table(file="genesXcells_introns_LLL91.tab",sep="\t",row.names=1,header=TRUE)
raw_counts92i<-read.table(file="genesXcells_introns_LLL92.tab",sep="\t",row.names=1,header=TRUE)
raw_counts93i<-read.table(file="genesXcells_introns_LLL93.tab",sep="\t",row.names=1,header=TRUE)
raw_counts94i<-read.table(file="genesXcells_introns_LLL94.tab",sep="\t",row.names=1,header=TRUE)
raw_counts96i<-read.table(file="genesXcells_introns_LLL96.tab",sep="\t",row.names=1,header=TRUE)
raw_counts97i<-read.table(file="genesXcells_introns_LLL97.tab",sep="\t",row.names=1,header=TRUE)
raw_counts98i<-read.table(file="genesXcells_introns_LLL98.tab",sep="\t",row.names=1,header=TRUE)
raw_counts99i<-read.table(file="genesXcells_introns_LLL99.tab",sep="\t",row.names=1,header=TRUE)
raw_counts100i<-read.table(file="genesXcells_introns_LLL100.tab",sep="\t",row.names=1,header=TRUE)
raw_counts101i<-read.table(file="genesXcells_introns_LLL101.tab",sep="\t",row.names=1,header=TRUE)
raw_counts102i<-read.table(file="genesXcells_introns_LLL102.tab",sep="\t",row.names=1,header=TRUE)



raw_counts_e<-cbind(raw_counts90e,raw_counts91e,raw_counts92e,raw_counts93e,raw_counts94e,
                    raw_counts96e,raw_counts97e,raw_counts98e,raw_counts99e,raw_counts100e,raw_counts101e,raw_counts102e)



raw_counts_i<-cbind(raw_counts90i,raw_counts91i,raw_counts92i,raw_counts93i,raw_counts94i,
                    raw_counts96i,raw_counts97i,raw_counts98i,raw_counts99i,raw_counts100i,raw_counts101i,raw_counts102i)

```




```{r}
# create a new variable from the rownames


# add rownames as a column in each data.frame and bind rows
counts<-bind_rows(raw_counts_e %>% add_rownames(), 
          raw_counts_i %>% add_rownames()) %>% 
    # evaluate following calls for each value in the rowname column
    group_by(rowname) %>% 
    # add all non-grouping variables
    summarise_all(sum)
rm(list=setdiff(ls(), "counts"))


counts<-as.data.frame(counts)
rownames(counts)<-counts[,'rowname']
badrows=grep("__",rownames(counts),value=T)
counts<-counts[!rownames(counts) %in% badrows,2:dim(counts)[2]]


```




replace known ENSG genes with their names, keep the remainder
```{r}
#load conversion table
conversiontable=read.csv("~/postdoc2/DCN_sequencing/human/EnsemblID_to_GeneSymbol.txt")
conversiontable.df<-as.data.frame(conversiontable)

conversiontable.notrepeated<-conversiontable[!(duplicated(conversiontable[,2])|duplicated(conversiontable[,2], fromLast=TRUE)),]
rownames(conversiontable.notrepeated)<-conversiontable.notrepeated[,1]


counts_translated<-counts[rownames(counts) %in% rownames(conversiontable.notrepeated),]


rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.name[match(rownames(counts_translated),conversiontable.notrepeated$Gene.stable.ID)])


counts_nottranslated<-counts[!rownames(counts) %in% rownames(conversiontable.notrepeated),]

counts2<-rbind(counts_translated,counts_nottranslated)
rm(list=setdiff(ls(), "counts2"))

```



```{r}

N49<- CreateSeuratObject(counts2, min.cells = 0, project = "DCN_plateseq")
#ERCC <- grep(pattern = "^ERCC", x = rownames(x = N49@assays$RNA@data), value = TRUE)

#N49 <- AddMetaData(object = N49, metadata = colSums(N49@assays$RNA@data[ERCC,])/colSums(N49@assays$RNA@data), col.name = "ERCC")
VlnPlot(N49,c("nFeature_RNA","nCount_RNA"))

```


annotate by donor: 
B1=18-T2789 white male
B2=19-T0161 white male
B3=19-T0469 (aka 449) black female
B4=18-T2807 white female
B5=?

```{r}
plates=sapply(strsplit(rownames(N49@meta.data),"_"),function(x) x[1])
B0=plates %in% c()
B1=plates %in% c("LLL90","LLL91","LLL92","LLL93","LLL97",
                 "LLL100","LLL101")
B2=plates %in% c()
B3=plates %in% c("LLL96","LLL98","LLL99","LLL102")
B4=plates %in% c()
B5=plates %in% c("LLL94")
donor=as.factor(as.numeric(B0)+2*as.numeric(B1)+3*as.numeric(B2)+4*as.numeric(B3)+5*as.numeric(B4)+6*as.numeric(B5))
#levels(donor)<-c("B0","B1","B2","B3","B4","B5")
levels(donor)<-c("B1","B3","B5")

N49$donor<-donor
```

annotate by sex

```{r}
sex=as.factor(donor %in% c("B3","B4"))
levels(sex)<-c('male','female')
```

annotate by FACS round:
```{r}
plates=sapply(strsplit(rownames(N49@meta.data),"_"),function(x) x[1])
R1=plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL51")
R2=plates %in% c("LLL50")
R3=plates %in% c("LLL51","LLL61","LLL53","LLL59","LLL60","LLL63")
R4=plates %in% c("LLL54","LLL55","LLL56")
R5=plates %in% c("LLL57","LLL58")
R6=plates %in% c("LLL90","LLL91","LLL92","LLL93","LLL94")
R7=plates %in% c("LLL96","LLL97","LLL98","LLL99","LLL100","LLL101","LLL102")
FACS=as.factor(as.numeric(R1)+2*as.numeric(R2)+3*as.numeric(R3)+4*as.numeric(R4)+5*as.numeric(R5)+6*as.numeric(R6)+7*as.numeric(R7))
N49$FACS<-FACS
```



annotate by DCN
```{r}
wells=sapply(strsplit(colnames(N49),"_"),function(x) x[2])
columns=as.numeric(substring(wells,2))
FNcells=(columns %in% c(1:12) & plates %in% c("LLL98")) |
  (columns %in% c(1:6) & plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL50","LLL51")) |
  plates %in% c("LLL60")
INcells=(columns %in% c(13:24)& plates %in% c("LLL98")) |
  plates %in% c("LLL92","LLL93","LLL82","LLL84")

DNdcells=(columns %in% c(13:18)& plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL50","LLL51","LLL59","LLL63")) |
  (columns %in% c(1:12) & plates %in% c("LLL52","LLL53","LLL57","LLL58","LLL61","LLL62")) 
DNvcells=(columns %in% c(19:24)& plates %in% c("LLL46","LLL47","LLL48","LLL49","LLL50","LLL51","LLL59","LLL63")) |
  (columns %in% c(13:24) & plates %in% c("LLL52","LLL53","LLL57","LLL58","LLL61","LLL62")) |
  (columns %in% c(17:24) & plates %in% c("LLL55","LLL56")) 
DNcells=plates %in% c("LLL29","LLL30","LLL31","LLL33",
                      "LLL79","LLL83")
FN_INcells= plates %in% c("LLL90","LLL91","LLL94","LLL96","LLL97","LLL99","LLL100","LLL101","LLL102")


DCN=as.factor(as.numeric(FNcells)+2*as.numeric(INcells)+3*as.numeric(DNdcells)+4*as.numeric(DNvcells)+5*as.numeric(DNcells)+6*as.numeric(FN_INcells))

N49@meta.data$DCN<-DCN


VlnPlot(object = N49, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)

```



```{r}
N49 <- NormalizeData(object = N49, normalization.method = "LogNormalize", scale.factor = 1e6)
N49<- subset(x = N49, subset = nFeature_RNA > 2000)
N49 <- FindVariableFeatures(object = N49,selection.method = "vst",nfeatures = 2000,verbose=FALSE)
N49 <- ScaleData(object = N49,vars.to.regress = c('nCount_RNA'))
N49 <- RunPCA(object = N49,npcs = 30,verbose = FALSE)
```

```{r,fig.width=15,fig.height=5}
VlnPlot(object = N49, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
```


```{r, fig.height=6, fig.width=10, warning=FALSE}
#N49 <- JackStraw(object = N49, dims=20)
#N49<- ScoreJackStraw(N49,dims=1:20)
#JackStrawPlot(object = N49,dims=1:20)
```

```{r, fig.height=6, fig.width=10, warning=FALSE}
ElbowPlot(object = N49)
```


```{r, fig.height=5, fig.width=7}
N49<- FindNeighbors(N49,dims=1:20)
N49 <- FindClusters(object = N49, resolution = 1)
```



```{r, fig.height=10, fig.width=15, message=FALSE, warnings=FALSE}
N49 <- RunTSNE(object = N49, dims = 1:20, perplexity=30, dim.embed = 2)

p1<-DimPlot(object = N49, reduction = 'tsne', label=TRUE)
p2<-DimPlot(N49,group.by = 'donor')
p3<-DimPlot(N49,group.by = 'DCN')
p4<-DimPlot(N49,group.by = 'orig.ident')
plot_grid(p1,p2,p3,p4,ncol=2)
```



```{r, fig.height=10, fig.width=15}
FeaturePlot(object = N49, features = c("SNAP25","SLC17A6","SLC6A5","GAD1","ETV1","SLC17A7","ALDOC"), reduction= "tsne")
```
cluster 14 is odd, but keep for now? 

```{r}
VlnPlot(object = N49, feature = c("nFeature_RNA"))
```
```{r}
VlnPlot(object = N49, feature = c("nCount_RNA"),pt.size = 0)

```




non-neuronal cells:
pdgfra(OPC),slc14A1(astrocyte),Gfap(astrocyte),Opalin(oligodendrocyte),nostrin(endothelial),tyrobp(microglia),slc1a3
```{r, fig.height=10, fig.width=15}
FeaturePlot(object = N49, features = c("OPC","SLC14A1","GFAP","OPALIN","NOSTRIN","TYROBP","SLC1A3"), reduction= "tsne")
```
```{r, fig.height=10, fig.width=15}
FeaturePlot(object = N49, features = c("SNAP25","MEG3","RBFOX3","SLC17A6","SLC17A7","SLC17A8","GAD1","GAD2","SLC6A5","SLC31A1","CALB1","CALB2"), reduction= "tsne")
```




split off neuronal cells

```{r}
neurons<-subset(x=N49,idents=c(2,6,5,1,0,3,9,14))
```

```{r}
neurons <- FindVariableFeatures(object = neurons,selection.method = "vst",nfeatures = 2000,verbose=FALSE)
neurons <- ScaleData(object = neurons,vars.to.regress = c('nCount_RNA','orig.ident','donor'))
neurons <- RunPCA(object = neurons,npcs = 30,verbose = FALSE)
```
```{r, fig.height=6, fig.width=10, warning=FALSE}
ElbowPlot(object = neurons)
```


```{r, fig.height=5, fig.width=7}
neurons<- FindNeighbors(neurons,dims=1:11)
neurons <- FindClusters(object = neurons, resolution = 1)
```



```{r, fig.height=10, fig.width=15, message=FALSE, warnings=FALSE}
neurons <- RunTSNE(object = neurons, dims = 1:11, perplexity=30, dim.embed = 2)


p1<-DimPlot(object = neurons, reduction = 'tsne', label=TRUE)
p2<-DimPlot(neurons,group.by = 'donor')
p3<-DimPlot(neurons,group.by = 'DCN')
p4<-DimPlot(neurons,group.by = 'orig.ident')
plot_grid(p1,p2,p3,p4,ncol=2)
```

meg3,slc17a6,slc6a5,gad1
```{r, fig.height=10, fig.width=15}
FeaturePlot(object = neurons, features = c("SNAP25","SLC17A6","SLC6A5","GAD1","TBR1","GRM2"), reduction= "tsne")

```
```{r}
markers<-FindAllMarkers(neurons,logfc.threshold = log(2))
markers %>% group_by(cluster) %>% top_n(5,avg_logFC)
```

save out neurons
```{r}
save(neurons,markers,file='neurons.RData')
```


