---
title: "Seurat alignment of excitatory mouse, human, and chicken cell types."
output: 
  pdf_document: 
    keep_tex: yes
keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Post novaseq49

```{r}
library(Seurat)
library(cowplot)
library(dplyr)
library(qlcMatrix)
library(Hmisc)
library(ComplexHeatmap)
library(circlize)
```

load mouse data


load mouse data, keeping all clusters 
```{r}
load(file = '~/postdoc2/DCN_sequencing/scRNAseq/plateseq/Novaseq31/dataforplotting.RData')
Idents(RL)<-'classes'
RL<-RenameIdents(RL,'ClassA'='Class-A','ClassB'='Class-B')
RL$classes<-Idents(RL)
RL$class<-paste0('M_',RL$classes)

mouse<-RL

rm(RL,VZ,merged)
mouse.data<-mouse@assays$RNA@counts
```

revert gene names to  ENSG genes 
```{r}
#load conversion table
conversiontable=read.csv("C:/Users/justus/Documents/postdoc2/DCN_sequencing/scRNAseq/plateseq/EnseblID_to_GeneSymbol.txt")
conversiontable.df<-as.data.frame(conversiontable)

conversiontable.notrepeated<-conversiontable[!(duplicated(conversiontable[,2])|duplicated(conversiontable[,2], fromLast=TRUE)),]
rownames(conversiontable.notrepeated)<-conversiontable.notrepeated[,1]


counts_translated<-mouse.data[rownames(mouse.data) %in% conversiontable.notrepeated[,2],]


rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.stable.ID[match(rownames(counts_translated),conversiontable.notrepeated$Gene.name)])


counts_nottranslated<-mouse.data[!rownames(mouse.data) %in% conversiontable.notrepeated[,2],]

mouse.data2<-rbind(counts_translated,counts_nottranslated)

```






load chickendata 
```{r}
load('~/postdoc2/DCN_sequencing/chicken/plateseq/novaseq44/ex.RData')
#define classes:
Idents(ex)<-'merged.res.3.m.renamed'
ex<-RenameIdents(ex,'Med.A1'="Class-A","MedL.A1"="Class-A","IntP.A1"="Class-A", "IntP.A2"="Class-A", "IntX.A1"="Class-A", "IntX.A2"="Class-A", "IntX.A3"="Class-A", "Med.B1"="Class-B",  "Med.B2"="Class-B",  "Med.B3"="Class-B",  "Med.B4"="Class-B","Med.B5"="Class-B",  "MedL.B1"="Class-B", "IntP.B1"="Class-B", "IntP.B2"="Class-B", "IntX.B1"="Class-B", "IntX.B2"="Class-B")
ex$classes<-Idents(ex)
ex$class<-paste0('C_',ex$classes)

all<-ex
```


```{r}
chicken.data<-all@assays$RNA@counts
chicken<-all
rm(ex_clean)
```


revert gene names to  ENSG genes 
```{r}
#load conversion table
conversiontable=read.csv("C:/Users/justus/Documents/postdoc2/DCN_sequencing/chicken/plateseq/EnsembleID99_to_GeneSymbol.txt")
conversiontable.df<-as.data.frame(conversiontable)

conversiontable.notrepeated<-conversiontable[!(duplicated(conversiontable[,2])|duplicated(conversiontable[,2], fromLast=TRUE)),]
rownames(conversiontable.notrepeated)<-conversiontable.notrepeated[,1]


counts_translated<-chicken.data[rownames(chicken.data) %in% conversiontable.notrepeated[,2],]


rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.stable.ID[match(rownames(counts_translated),conversiontable.notrepeated$Gene.name)])


counts_nottranslated<-chicken.data[!rownames(chicken.data) %in% conversiontable.notrepeated[,2],]

chicken.data2<-rbind(counts_translated,counts_nottranslated)

```

load human data

```{r}
load('~/postdoc2/DCN_sequencing/human/plateseq/Novaseq49/dataforplotting.RData')
Idents(ex)<-'merged.res.2.renamed'
ex<-RenameIdents(ex,"Med.A1" = 'Class-A', "MedL.A1" = 'Class-A', "MedL.A2" = 'Class-A', "IntA.A1" = 'Class-A', "IntP.A1" = 'Class-A', "IntP.A2" = 'Class-A', "Med.B1" = 'Class-B',  "IntA.B1" = 'Class-B', "IntP.B1" = 'Class-B', "IntP.B2" = 'Class-B', "Lat.B1" = 'Class-B')
ex$classes<-Idents(ex)
ex$class<-paste0('H_',ex$classes)

all<-ex
rm(ex,inh)

human.data<-all@assays$RNA@counts
human<-all
```


revert gene names to  ENSG genes 
```{r}
#load conversion table
conversiontable=read.csv("C:/Users/justus/Documents/postdoc2/DCN_sequencing/human/EnsembleID_to_GeneSymbol.txt")
conversiontable.df<-as.data.frame(conversiontable)

conversiontable.notrepeated<-conversiontable[!(duplicated(conversiontable[,2])|duplicated(conversiontable[,2], fromLast=TRUE)),]
rownames(conversiontable.notrepeated)<-conversiontable.notrepeated[,1]


counts_translated<-human.data[rownames(human.data) %in% conversiontable.notrepeated[,2],]


rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.stable.ID[match(rownames(counts_translated),conversiontable.notrepeated$Gene.name)])


counts_nottranslated<-human.data[!rownames(human.data) %in% conversiontable.notrepeated[,2],]

human.data2<-rbind(counts_translated,counts_nottranslated)

```












read in ortholog relationships obtained from ensemble biomart

```{r}
orth.table<-read.csv(file='~/postdoc2/DCN_sequencing/chicken/plateseq/chicken_human_mouse_orthologs.txt',header=T)
#throw out anything that isnt onetoone
orth.table<-orth.table[orth.table[,"Human.homology.type"]=="ortholog_one2one" & orth.table[,"Mouse.homology.type"]=="ortholog_one2one",]
#drop extra columns
orth.table.clean<-unique(orth.table[,c("Gene.stable.ID",'Human.gene.stable.ID','Human.gene.name','Mouse.gene.stable.ID','Mouse.gene.name')])
```


throw out what isn't in a 1:1 ortholog
```{r}
chicken.data.filtered<-chicken.data2[rownames(chicken.data2) %in% as.character(orth.table.clean$Gene.stable.ID),]
rownames(chicken.data.filtered)<-as.character(orth.table.clean$Mouse.gene.stable.ID[match(rownames(chicken.data.filtered),orth.table.clean$Gene.stable.ID)])

human.data.filtered<-human.data2[rownames(human.data2) %in% as.character(orth.table.clean$Human.gene.stable.ID),]
rownames(human.data.filtered)<-as.character(orth.table.clean$Mouse.gene.stable.ID[match(rownames(human.data.filtered),orth.table.clean$Human.gene.stable.ID)])

mouse.data.filtered<-mouse.data2[rownames(mouse.data2) %in% as.character(orth.table.clean$Mouse.gene.stable.ID),]
rownames(mouse.data.filtered)<-as.character(orth.table.clean$Mouse.gene.stable.ID[match(rownames(mouse.data.filtered),orth.table.clean$Mouse.gene.stable.ID)])

```

why are there still different numbers of genes?
```{r}
chicken.data.filtered2<-chicken.data.filtered[rownames(chicken.data.filtered) %in% rownames(human.data.filtered) & rownames(chicken.data.filtered) %in% rownames(mouse.data.filtered),]
human.data.filtered2<-human.data.filtered[rownames(human.data.filtered) %in% rownames(chicken.data.filtered2) & rownames(human.data.filtered) %in% rownames(mouse.data.filtered),]

mouse.data.filtered2<-mouse.data.filtered[rownames(mouse.data.filtered) %in% rownames(chicken.data.filtered2) & rownames(mouse.data.filtered) %in% rownames(human.data.filtered2),]
```

now convert them all back to human names for simplicity.

replace known ENSMUSG genes with their names, keep the remainder
```{r}
#load conversion table
conversiontable=read.csv("C:/Users/justus/Documents/postdoc2/DCN_sequencing/scRNAseq/plateseq/EnseblID_to_GeneSymbol.txt")
conversiontable.df<-as.data.frame(conversiontable)

conversiontable.notrepeated<-conversiontable[!(duplicated(conversiontable[,2])|duplicated(conversiontable[,2], fromLast=TRUE)),]
rownames(conversiontable.notrepeated)<-conversiontable.notrepeated[,1]


counts_translated<-chicken.data.filtered2[rownames(chicken.data.filtered2) %in% rownames(conversiontable.notrepeated),]
rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.name[match(rownames(counts_translated),conversiontable.notrepeated$Gene.stable.ID)])
counts_nottranslated<-chicken.data.filtered2[!rownames(chicken.data.filtered2) %in% rownames(conversiontable.notrepeated),]
chicken.data.filtered3<-rbind(counts_translated,counts_nottranslated)


counts_translated<-human.data.filtered2[rownames(human.data.filtered2) %in% rownames(conversiontable.notrepeated),]
rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.name[match(rownames(counts_translated),conversiontable.notrepeated$Gene.stable.ID)])
counts_nottranslated<-human.data.filtered2[!rownames(human.data.filtered2) %in% rownames(conversiontable.notrepeated),]
human.data.filtered3<-rbind(counts_translated,counts_nottranslated)

counts_translated<-mouse.data.filtered2[rownames(mouse.data.filtered2) %in% rownames(conversiontable.notrepeated),]
rownames(counts_translated)<-as.character(conversiontable.notrepeated$Gene.name[match(rownames(counts_translated),conversiontable.notrepeated$Gene.stable.ID)])
counts_nottranslated<-mouse.data.filtered2[!rownames(mouse.data.filtered2) %in% rownames(conversiontable.notrepeated),]
mouse.data.filtered3<-rbind(counts_translated,counts_nottranslated)



```






Set up seurat objects. 

```{r}
human2<-CreateSeuratObject(human.data.filtered3)
human2@meta.data<-human@meta.data
human2$species<-'human'
human2$clusters<-paste0('H_',human2$merged.res.2.renamed)
human2$clusters<-factor(human2$clusters,levels=paste0('H_',levels(human2$merged.res.2.renamed)))


chicken2<-CreateSeuratObject(chicken.data.filtered3)
chicken2@meta.data<-chicken@meta.data
chicken2$species<-'chicken'
chicken2$clusters<-paste0('C_',chicken2$merged.res.3.m.renamed)
chicken2$clusters<-factor(chicken2$clusters,levels=paste0('C_',levels(chicken2$merged.res.3.m.renamed)))


mouse2<-CreateSeuratObject(mouse.data.filtered3)
mouse2@meta.data<-mouse@meta.data
mouse2$species<-'mouse'
mouse2$clusters<-paste0('M_',mouse2$final.clusters2)
mouse2$clusters<-factor(mouse2$clusters,levels=paste0('M_',levels(mouse2$final.clusters2)))

dcn<-merge(mouse2,chicken2)
dcn<-merge(dcn,human2)

#downsample to roughly equal cluster sizes
#Idents(dcn)<-'clusters'
#dcn<-subset(dcn, downsample = 50)


load('~/postdoc2/DCN_sequencing/human/plateseq/Novaseq49/3species_downsampled50.RData')


dcn.list<-SplitObject(dcn,split.by = 'species')


for (i in 1:length(x = dcn.list)) {
    dcn.list[[i]] <- NormalizeData(object = dcn.list[[i]], verbose = FALSE,scale.factor = 1e6)
    dcn.list[[i]] <- FindVariableFeatures(object = dcn.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
library(future)
plan("multiprocess", workers = 4)
plan("multiprocess",workers=4)
options(future.globals.maxSize=2971520000)

DCN.anchors <- FindIntegrationAnchors(object.list = dcn.list, dims = 1:40,anchor.features = 2000)
```

```{r}
DCN.integrated <- IntegrateData(anchorset = DCN.anchors, dims = 1:40,features.to.integrate = rownames(dcn))
```

```{r}
library(ggplot2)
library(cowplot)
DefaultAssay(object = DCN.integrated) <- "integrated"

DCN.integrated <- ScaleData(object = DCN.integrated, verbose = FALSE,vars.to.regress = c())
DCN.integrated<-RunPCA(DCN.integrated,dims=1:40)
ElbowPlot(DCN.integrated)
```

```{r}
useful.dims=c(1:5)
DCN.integrated<-FindNeighbors(DCN.integrated,dims = useful.dims)
DCN.integrated<-FindClusters(DCN.integrated,resolution = 0.2)
DCN.integrated <- RunTSNE(object = DCN.integrated, reduction = "pca", 
    dims = useful.dims,perplexity=50)

```
```{r,fig.width=15,fig.height=10}
p1<-DimPlot(DCN.integrated,group.by = 'species',label = F,pt.size = 1.5,cols=c('gold','maroon','grey'))+NoLegend()
p2<-DimPlot(DCN.integrated,group.by = 'classes',label=F,pt.size = 1.5,cols=c('#ff6347','#00688b'))+NoLegend()
p3<-DimPlot(DCN.integrated,label=F,pt.size = 1.5)+NoLegend()
p4<-DimPlot(DCN.integrated,group.by = 'clusters',label=T,pt.size = 1.5)+NoLegend()
plot_grid(p1,p2,p3,p4,ncol=2)
```








```{r,fig.height=2.5,fig.width=10}
t<-table(DCN.integrated$integrated_snn_res.0.2,DCN.integrated$class)
t2<-prop.table(t,2)
mat<-matrix(as.numeric(data.matrix(t2)),nrow = dim(t2)[1],ncol=dim(t2)[2],dimnames = list(rownames(t2),colnames(t2)))
col1=colorRamp2(c(0,0.5, 1), c("navy", "white", "firebrick3"),space="sRGB")


#column annotation
#ha_bottom<- HeatmapAnnotation(
#  species=colnames(mat),
#  col=list(species=setNames(c(rep('gold',17),rep('grey50',15)),colnames(mat))),
#  simple_anno_size=unit(2,"mm")
#)


Heatmap(mat,col=col1,cluster_columns = F,cluster_rows = F)
```
